Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-10-25 GWLDCII.tex
talking on the phone

In recent years, there has been significant interest in understanding the geometry and combinatorics underlying the field theory SYM $N=4$ \cite{wilsonloop, Arkani-Hamed:2013jha, Amplituhedronsquared, galashinlam, AmplituhedronDecomposition}. This started with the observation that BCFW diagrams, which represent the on shell interactions of this theory, correspond to plabic graphs \cite{GrassmannAmplitudebook}. Plabic graphs were introduced by Postnikov in \cite{Postnikov}, and provide one of the many equivalent ways to classify the positroid cells of the positive real Grassmannian $\Gr(k,n)$. This led to the representation of the on shell integrals associated to this theory in terms of a geometric space (the Amplituhedron) embedded in $\Gr(k,n)$ \cite{Arkani-Hamed:2013kca,Arkani-Hamed:2013jha}. Moreover, integrals associated to the BCFW diagrams define a volume form on the Amplituhedron. This gives a novel and elegant interpretation of scattering amplitudes for this theory, as well as a means of greatly simplifying the calculations involved in traditional Feynman diagram approaches. Since then, there has been significant work to understand the volume form and its physical implications \cite{UnwindingAmplituhedron,galashinlam}.

Meanwhile, a different body of work has studied SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv,Adamo:2012xe,  Boels:2007qn, Bullimore:2010pj}. These integrals are calculated via holomorphic Wilson loops. As with the Amplituhedron, each Wilson loop represents the sum of a family of $N^kMHV$ Feynman diagrams, and adding up the corresponding sums of Feynman integrals associates a volume form to a convex space different from the Amplituhedron \cite{Hodges2013}. In  \cite{wilsonloop}, Agarwala and Marin-Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$. Other work has shown a connection between the volume filled by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}. Specifically, Marcott shows that the space parametrized by the Wilson loop diagrams is  exactly the correct positroid cell in $\Gr(k,n)$, up to a set of measure $0$ \cite{WLDdim}.

@@

This paper continues this body of work by establishing concrete connections between the combinatorics of the Wilson loop diagram as determined by the physics, and the combinatorics of the associated positroid cell as identified in \cite{wilsonloop}. In this paper, we give an algorithm to read the Grassmann necklace, one of the standard combinatorial characterizations of the positroid cell, directly off the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. This apparent contradiction led to the work in \cite{non-orientable}, which showed that the actual geometric space defined by the Wilson loop diagrams is a non-orientable $k$-dimensional bundle over this $3k$-dimensional manifold in $\Gr(k,n)$.

>>>>

In recent years, there has been significant interest in understanding the geometry and combinatorics underlying the field theory SYM $N=4$ \cite{wilsonloop, Arkani-Hamed:2013jha, Amplituhedronsquared, galashinlam, AmplituhedronDecomposition}. This started with the observation that BCFW diagrams, which represent the on shell interactions of this theory, correspond to plabic graphs \cite{GrassmannAmplitudebook}. Plabic graphs were introduced by Postnikov in \cite{Postnikov}, and provide one of the many equivalent ways to classify the positroid cells of the positive real Grassmannian $\Gr(k,n)$. This led to the representation of the on shell integrals associated to this theory in terms of a geometric space (the Amplituhedron) embedded in $\Gr(k,n)$ \cite{Arkani-Hamed:2013kca,Arkani-Hamed:2013jha}. Moreover, integrands associated to the BCFW diagrams define a volume form on the Amplituhedron. This gives a novel and elegant interpretation of scattering amplitudes for this theory, as well as a means of greatly simplifying the calculations involved in traditional Feynman diagram approaches. Since then, there has been significant work to understand the volume form and its physical implications \cite{UnwindingAmplituhedron,galashinlam}.

Meanwhile, a different body of work has studied SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv,Adamo:2012xe,  Boels:2007qn, Bullimore:2010pj}. These integrals are calculated via holomorphic Wilson loops. As with the Amplituhedron, each Wilson loop represents the sum of a family of $N^kMHV$ Feynman diagrams, and adding up the corresponding sums of Feynman integrands associates a volume form to a convex space different from those in the decomposition the Amplituhedron \cite{Hodges2013}. In  \cite{wilsonloop}, Agarwala and Marin-Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$. Other work has shown a connection between the volume filled by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}. Specifically, Marcott shows that the space parametrized by the Wilson loop diagrams is  exactly the correct positroid cell in $\Gr(k,n)$, up to a set of measure zero \cite{WLDdim}.

@@

This paper continues this body of work by establishing concrete connections between the combinatorics of the Wilson loop diagram as determined by the physics, and the combinatorics of the associated positroid cell as identified in \cite{wilsonloop}. In this paper, we give an algorithm to read the Grassmann necklace, one of the standard combinatorial characterizations of the positroid cell, directly off the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. This apparent contradiction is resolved by the non-orientability result of \cite{non-orientable}.

%As described above, each integrand associated to a Wilson loop diagram gives the volume of a convex polytope.  These individual volumes are not finite (*or maybe it needs to say "not necessarily finite" or "not always finite" or whatever is true*) which corresponds to the fact that the Feynman integrals have singularities coming from factors in their denominators.  However SYM $N=4$ is a finite theory \cite{Adamo:2011pv,Adamo:2012xe,Arkani-Hamed:2013jha}, so the singularities must cancel out in the sum over all diagrams.  This should mean that when the polytopes for each Wilson loop diagram are put together into one object (*you said union, but I think it must be more complicated than that, a set theoretic union can't make the infinite finite*) the overall volume is finite.  This cancellation has been conjectured in general \cite{Hodges2013} and verified explicitly in a few cases \cite{casestudy, HeslopStewart}, but it is hard to prove in general due to the fact that the relationship between Wilson loop diagrams and $3k$-dimensional positroid cells of $\Gr(k,n)$ is neither one-to-one nor onto.

%In this paper, we give a geometric characterization of the factors appearing in the denominator of these Feynman integrals in terms of the Grassmann necklace of the associated positroid. Specifically, we show that the set of irreducible factors in the denominator, as defined in the physics literature \cite{HeslopStewart}, are exactly the irreducible factors of the minors defined by the Grassmann necklace evaluated on a physically significant realization of the associated positroid cell. In other words, the singularities arising from tree level particle interactions in SYM $N=4$ theory do not just come from the Feynman rules describing the theory: they can also be seen directly from the associated geometry, given the physically correct representation.  We hope these results will pave the way for a general study of these poles in future.
